# (10 points) ARIMA model

Consider `fma::sheep`, the sheep population of England and Wales from 1867--1939. :sheep: 

```{r load sheep data, message=FALSE, warning=FALSE}
#install.packages('fma')
library(fma)
sheep_ts <- as_tsibble(fma::sheep)
head(sheep_ts)
```

## Time series plot
Produce a time plot of the time series, and comment on what you observe.

```{r produce a time series plot of sheep}
sheep_plot <- autoplot(sheep_ts) + labs(x = "Year", y = TeX(r'(Sheep $\times 10^{-6}$)'))
sheep_plot
```

> Sheep population is gradually decreasing over this period and appears to stabilize in the 1930's at around 1700$\times 10^{6}$. There is a significant local dip that coincides with the WWI and subsequent period of economic turmoil in Europe. 

## Fit a model 

Assume you decide to fit the following model: 

$$
  y_{t} = y_{t-1} + \phi_1(y_{t-1}-y_{t-2}) + \phi_2(y_{t-2}-y_{t-3}) + \phi_3(y_{t-3}-y_{t-4}) + \epsilon_t,
$$ 

where $\epsilon_t$ is a white noise series.

### Model type
What sort of ARIMA model is this (i.e., what are $p$, $d$, and $q$)?

> Thi is a representation of ARIMA(3.1.0). There are no error terms from the past lags, therefore MA term q = 0; there is only one differing step, so d = 1, and the last lag taken into account is $y_{t-4}$ which means p = 3 (given d = 1).

### Back to the future 

Express this ARIMA model using backshift operator notation.

> $(1-B)y_{t} = \phi_1(1-B)y_{t-1} + \phi_2(1-B)y_{t-2} + \phi_3(1-B)y_{t-3} + \epsilon_t$

> $(1-B)y_{t} = \phi_1(1-B)By_{t} + \phi_2(1-B)B^2y_{t} + \phi_3(1-B)B^3y_{t} + \epsilon_t$

> $(1-B)y_{t} = (1-B)(\phi_1By_{t} + \phi_2B^2y_{t} + \phi_3B^3y_{t}) + \epsilon_t$

> $(1-B)(1-\phi_1B-\phi_2B^2-\phi_3B^3)y_{t} = \epsilon_t$

## Is this model appropriate? 

Examine the ACF and PACF of the differenced data. Evaluate whether this model is appropriate. 

```{r difference the data}
# Add a column with differenced data to the original tsibble
sheep_ts <- sheep_ts %>% mutate(diff = difference(value, lag = 1))
# Generate acf plot
acf_plot <- acf(sheep_ts$diff, na.action = na.pass, plot = FALSE) %>%
  autoplot() + labs(title = "ACF for differenced data")
# Generate pacf plot
pacf_plot <- pacf(sheep_ts$diff, na.action = na.pass, plot = FALSE) %>%
  autoplot() + labs(title = "PACF for differenced data")
acf_plot | pacf_plot # Display plots
```

> The model does not capture all the aspects of the underlying data. On one hand, PACF plot shows only 3 signofocant lags, indicating that thsi si indeed AR(3) process. On the other hand, ACF plot shows abrupt decrease of significance after lag 4 and some oscillating pattern, indicating MA(4) process with some periodicity. Therefore ARIMA(3.1.4) would be more appropriate model.

## Forecasts, by hand! 
The last five values of the series are given below:

| Year              | 1935 | 1936 | 1937 | 1938 | 1939 |
|:------------------|-----:|-----:|-----:|-----:|-----:|
| Millions of sheep | 1648 | 1665 | 1627 | 1791 | 1797 |

The estimated parameters are: 

- $\phi_1 =  0.42$; 
- $\phi_2 = -0.20$;  and, 
- $\phi_3 = -0.30$.

Without using the forecast function, calculate forecasts for the next three years (1940--1942).

$y_{1940} = y_{1939} + \phi_1(y_{1939}-y_{1938}) + \phi_2(y_{1938}-y_{1937}) + \phi_3(y_{1937}-y_{1936}) + 0$

$y_{1941} = y_{1940} + \phi_1(y_{1940}-y_{1939}) + \phi_2(y_{1939}-y_{1939}) + \phi_3(y_{1938}-y_{1937}) + 0$

$y_{1942} = y_{1941} + \phi_1(y_{1941}-y_{1940}) + \phi_2(y_{1940}-y_{1939}) + \phi_3(y_{1939}-y_{1938}) + 0$


```{r create forcasts by hand}
y1936<-1665
y1937<-1627 
y1938<-1791
y1939<-1797
phi_1 <-  0.42
phi_2 <- -0.20
phi_3 <- -0.30

y1940 <- y1939 + phi_1*(y1939-y1938) + phi_2*(y1938-y1937) + phi_3*(y1937-y1936)
y1941 <- y1940 + phi_1*(y1940-y1939) + phi_2*(y1939-y1939) + phi_3*(y1938-y1937)
y1942 <- y1941 + phi_1*(y1941-y1940) + phi_2*(y1940-y1939) + phi_3*(y1939-y1938)
```

> Forecasted values for 1940, 1941 and 1942 are `r round(y1940, digits = 0)`, `r round(y1941, digits = 0)` and `r round(y1942, digits = 0)`, respectively.

## Interpret roots
Find the roots of your model's characteristic equation. Is this process stationary?.

```{r find roots of the charachteristic polynomial}
roots <- polyroot(c(1, -phi_1, -phi_2, -phi_3))
mroot1 <- Mod(roots[1])
mroot2 <- Mod(roots[2])
mroot3 <- Mod(roots[3])
```
> Charachteristic polynomial for the ARMA model is $1-\phi_1B-\phi_2B^2-\phi_3B^3 = 0$. For the process to be stationary, all of the roots should lle strictly outside of a unit circle on the complex plane, i.e. modules of all roots should be more than 1. We have modules `r round(mroot1, digits =2)`, `r round(mroot2, digits =2)`, `r round(mroot3, digits =2)`, all above 1. That means that the process is stationary. 
