# (12 points) Vector autoregression

```{r load VAR packages, message=FALSE,message=FALSE}
library(tidyverse)
```

Annual values for real mortgage credit (RMC), real consumer credit (RCC) and real disposable personal income (RDPI) for the period 1946-2006 are recorded in `./data/mortgage_credit.csv`. 

All of the observations are measured in billions of dollars, after adjustment by the Consumer Price Index (CPI). 

Our goal is to develop a VAR model for these data for the period 1946-2003, and then forecast the last three years, 2004-2006. 


```{r read credit data, message=FALSE}
credit <- read_csv('./data/mortgage_credit.csv')
cred_ts <- as_tsibble(credit, index = Year)
#glimpse(credit)
```

## Time series plot

Plot the time-series of real mortgage credit (RMC), real consumer credit (RCC) and real disposable personal income (RDPI)? Do they look stationary?

```{r Time series plot}
#Fill this in

```

> 'Fill this in


## Check for the unit root

Plot ACF/PACF and Perform the unit root test on these variables and report the results. Do you reject the null of unit root for them? Is the first differencing necessary? 
```{r ACF/PACF and  unit root test}
cred_ts %>% gg_tsdisplay(RMC, plot_type="partial") +labs(subtitle = "RMC")
cred_ts %>% gg_tsdisplay(RCC, plot_type="partial") +labs(subtitle = "RCC")
cred_ts %>% gg_tsdisplay(RDPI, plot_type="partial") +labs(subtitle = "RDPI")

```


```{r ACF/PACF and  unit root test}
RMC_kpss <- cred_ts %>% features(RMC, unitroot_kpss)
RCC_kpss <- cred_ts %>% features(RCC, unitroot_kpss)
RDPI_kpss <- cred_ts %>% features(RDPI, unitroot_kpss)
```

The ACF oa all three series decay slowly, and The first lag of their PACF is 1 in all three cases, which suggests non-stationaruty. The KPSS tests for RMC, RCC and RDPI reject the null hypothesis with the p values of `r round(RMC_kpss$kpss_pvalue, digits = 3)`,  es should have a unit root. The KPSS tests for RMC, RCC and RDPI reject the null hypothesis with the p values of `r round(RMC_kpss$kpss_pvalue, digits = 3)`,`r round(RCC_kpss$kpss_pvalue, digits = 3)` and `r round(RDPI_kpss$kpss_pvalue, digits = 3)` respectively, suggesting that the series are not stationary.

d) Now calculate the first differences of the series
```{r}
cred_diff <- cred_ts %>% 
mutate(diff_RMC = difference(RMC), diff_RCC = difference(RCC), diff_RDPI = difference(RDPI))
RMC_kpss_diff <- cred_diff %>% features(diff_RMC, unitroot_kpss)
RCC_kpss_diff <- cred_diff %>% features(diff_RCC, unitroot_kpss)
RDPI_kpss_diff <- cred_diff %>% features(diff_RDPI, unitroot_kpss)
```

## Determine VAR model

Based on the unit root results transform the variables and determine the lag length of the VAR using the information criteria.

```{r Determine VAR model}
df1 <- df %>%
mutate(
diff_bitcoin = difference(bitcoin),
diff_google = difference(google)
) %>%
filter(Date >= "2020-01-12") %>%
dplyr::select(diff_bitcoin,diff_google)
VARselect(df1, lag.max = 4, type="none")
```

We use different model selection criteria to choose the optimal number of lags. AIC, HQ, and FPE suggest three lags while SC
(akin to BIC in the univariate case) suggests 1 lag. Here we estimate VAR(3)


## Estimation 

Estimate the selected VAR in previous part and comment on the results.

```{r Estimate VAR}
var_diff = vars::VAR(df1, p = 3, type = "none")
summary(var_diff)
```

In the first equation, none of the coefficients are statistically significant in 5% for change in bitcoin regression. R-squared is small,
and only 6.9 percent of the variations of change in bitcoin prices can be explained by the lagged change in google search volume
and lagged change in bitcoin prices.
In the second equation, for change in google trend, The estimated VAR model suggests that the past values of change in bitcoin
prices have explanatory power for current values of change in google trend.
However, we find that only lag one and three of the bitcoin prices is significant at a 10% percent significance level.
So apparently, large fluctuations in bitcoin prices lead to higher attention to the bitcoin and higher google search volume.
Also, 21 percent of the variations of change in google search volume can be explained by the lag of change in google search volume
and the lag of change in bitcoin prices.
So, based on these results, we can conclude that higher bitcoin prices could predict higher google search volume, but not the other
way around.


## Model diagnostic

Do diagnostic checking of the VAR model.

```{r Model diagnostic}
roots(var_diff)

var_diff_test<- serial.test(var_diff, lags.pt = 12)
var_diff_test

```

First, we need to check if the estimated VAR(3) is a stable process, and we will need to check if the eigenvalues of the companion
matrix are all less than one.
Here, since here all eigenvalues are less than 1, VAR(3) is a stable process.
Then, we need to check for tests for autocorrelation in residuals. The serial.test() computes the multivariate Portmanteau- for serial correlation.
Based on the test results, the null hypothesis of no autocorrelation is not rejected since the p-value is 0.4868.

## Forecasting

forecast the last three years, 2004-2006.


```{r Forecasting}
var_diff = vars::VAR(as.ts(df1), p = 3, type = "none")
forecast(var_diff) %>%
autoplot() +
xlab("Weeks")
```

> Both changes in log of bitcoin prices and google trend search revert to their means, which is onsistent with stationarity.
